\OurTool{}: Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles

### 2020-07-03 IPANEMAP-2019.tex
Updates from Overleaf

\noindent We introduce \OurTool{}, an automated method for the modeling of RNA structure from multiple probing reactivity profiles. Input profiles can result from experiments based on diverse protocols, reagents, \RevA{or collection of variants}, and are jointly analyzed to predict the dominant conformations of an RNA.

@@

To the best of our knowledge, \OurTool currently represents the only available method for a joint automated consideration of multiple probing profiles. 
%Surprisingly, it also appears to produce the most accurate secondary structure models when informed by a single condition, as demonstrated by our reanalysis on the Hajdin\etal\cite{Hajdin2013} benchmark. We attribute this unexpected result to the adoption of a sampling/clustering approach, \RevA{which appears to eliminate spurious outliers from the conformational space even when a single conformation dominates.}
%\RevA{In this context, the good performance of \OurTool in comparison with \Software{RSample}~\cite{Spasic2017} may appear surprising. Indeed, while independently developed and used~\cite{Deforges2017}, both methods share a common sampling/clustering scheme. However, \Software{RSample}~\cite{Spasic2017} reverts to a MEA~\cite{Lu2009} computation whenever a single conformation is suspected. The MEA structure coincides with the centroid of the whole Boltzmann ensemble, suggesting that the increased performances of \OurTool could be at least partially attributed to the use of sampling, and to our adaptive choice of the number of clusters.}
%Moreover, even in cases where a single dominant conformation is expected, considering multiple clusters seems to create more opportunities for discarding noisy random unstable conformations. Such conformations naturally occur in a Boltzmann-distributed subset of structures, and may end up \emph{polluting} the centroid structures by artificially inflating the cluster diversity. 
%The impact on \Software{RSample} predictions of such a \emph{noise} seems worsened by the fact that it returns the MEA, the centroid of the whole ensemble, whenever a dominant conformation is detected.
Our analyses reveal that integrating multiple profiles of at least three conditions allows to mitigate the effect of outlier conditions and, at times, to produce better secondary structure models than the one inferred only from the most accurate profile. A comparison of the joint performances to the average appears fair since no single reagent appears to systematically dominate in term of predictive capacity. It follows that a modeler cannot objectively favor \emph{a priori} a probing condition/reagent~\cite{Yu2018}. In other words, multiple experimental probing profiles, even combined in a pairwise fashion, \RevA{can be used jointly to mitigate the empirical risk of a misprediction}.

@@

\section*{Acknowledgements}
The authors wish to express their gratitude to Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble, Nathalie Chamond for helpful suggestions and Benoit Masquida for the gift of \didy{} plasmid and helpful discussions about protocols and \didy{} structure .

>>>>

\noindent We introduce \OurTool{}, \RevB{the first} automated method for the modeling of RNA structure from multiple probing reactivity profiles. Input profiles can result from experiments based on diverse protocols, reagents, \RevA{or collection of variants}, and are jointly analyzed to predict the dominant conformations of an RNA.

@@

Our analyses reveal that integrating multiple profiles of at least three conditions allows to mitigate the effect of outlier conditions and, at times, to produce better secondary structure models than the one inferred only from the most accurate profile. A comparison of the joint performances to the average appears fair since no single reagent appears to systematically dominate in term of predictive capacity. It follows that a modeler cannot objectively favor \emph{a priori} a probing condition/reagent~\cite{Yu2018}. In other words, multiple experimental probing profiles, even combined in a pairwise fashion, \RevA{can be used jointly to mitigate the empirical risk of a misprediction}.

To the best of our knowledge, \OurTool currently represents the only available method for a joint automated consideration of multiple probing profiles. \RevB{While hard constraints~\cite{Mathews2004} could in principle support multiple sources of probing data, their derivation from reactivities usually require a careful choice of cutoffs, making their automation virtually impossible. Overly liberal cutoffs will often induce contradictory constraints (i.e. no compatible structure) across conditions, while conservative cutoffs will lead to sparse constraints, practically wasting most of the probing-derived information. In fact the tediousness, and suboptimality, of the manual determination of those cutoffs while modeling HIV-1 structural elements~\citep{Deforges2017}, was one of the main motivation behind the present work.}

@@

\section*{Acknowledgements}
The authors wish to express their gratitude to: Ronny Lorenz for developing a probing-informed version of \Software{RNAsubopt}, allowing to sample the pseudo-Boltzmann ensemble; Nathalie Chamond for helpful suggestions; Benoit Masquida for the gift of \didy{} plasmid and helpful discussions about protocols and \didy{} structure; \RevB{and an anonymous reviewer for helping us detect an error in our execution of \Software{RSample}}.
